---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(ggplot2)
data <- read.csv("./data/activity.csv")
head(data)
```

## What is mean total number of steps taken per day?

* Total of activity by day:

```{r}
total <- aggregate(steps ~ date, data = data, FUN=sum)
total
```

```{r}

ggplot(total, aes(x=steps)) + 
  geom_histogram(binwidth=1000) + labs(title = "Histogram of the total number of steps taken each day", x="Steps", y="Frequency")

```


* Median
```{r}
median(total$steps)
```

* Mean
```{r}
mean(total$steps)
```

## What is the average daily activity pattern?

```{r}

totalInterval <- aggregate(steps ~ interval, data = data, FUN=sum)
ggplot(totalInterval, aes(x=interval, y=steps)) + geom_line() + labs(title = "Plot of the total steps taken interval of each date", x="Interval", y="Total steps")

```

* Best 5 minute interval:

```{r}
idx <- which(totalInterval$steps == max(totalInterval$steps))
totalInterval$interval[idx]
```

## Imputing missing values

* Total of missings:

```{r}
nrow(data[is.na(data), ])
```

* Imput missing values by mean:

```{r}

columns = c('steps', 'interval')
for(column in columns){
  data[is.na(data[,column]), column] <- mean(data[,column], na.rm = TRUE)
}

nrow(data[is.na(data), ])

```

```{r}
total <- aggregate(steps ~ date, data = data, FUN=sum)
ggplot(total, aes(x=steps)) + 
  geom_histogram(binwidth=1000) + labs(title = "Histogram of the total number of steps taken each day whithout NA", x="Steps", y="Frequency")
```

* Median
```{r}
median(total$steps)
```

* Mean
```{r}
mean(total$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

* Creating a new variable with type of day: 
```{r}

data$daytype <- ifelse(as.POSIXlt(data$date)$wday %in% c(6,0), 'Weekend', 'Weekday')

```


```{r}
totalInterval <- aggregate(steps ~ interval * daytype, data = data, FUN=sum)
ggplot(totalInterval, aes(x=interval, y=steps)) + geom_line() + facet_wrap(~ daytype) + labs(title = "Plot of the total steps taken interval of each date, sep. by type of day", x="Interval", y="Total steps")

```

Note that the interval between 800 and 1000 of weekday is much more intense compared by weekend.